---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
author: "Denitsa Panova"
date: "March 16, 2017"
---


## Loading and preprocessing the data
It is now possible to collect a large amount of data about personal movement using   
activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up.   
These type of devices are part of the “quantified self” movement – a group of   
enthusiasts who take measurements about themselves regularly to improve their   
health, to find patterns in their behavior, or because they are tech geeks.   
But these data remain under-utilized both because the raw data are hard to obtain   
and there is a lack of statistical methods and software for processing and interpreting the data.  
  
This assignment makes use of data from a personal activity monitoring device.  
This device collects data at 5 minute intervals through out the day.   
The data consists of two months of data from an anonymous individual collected   
during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.  

*Read in the data
``` {r echo=TRUE}
data = read.csv('activity.csv')
```

_Note that you need to position the markdown and the data in the same directory._
* Check what are the data type for each column 
```{r echo=TRUE}
str(data)
```
Everything looks ok with how thw data is read in. We note that there are 61 days.


## What is mean total number of steps taken per day?



## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
